---
title: "Reanalysis of 03-Zecca"
author: "Simon Schwab^[University of Zurich, simon.schwab@uzh.ch]"
date: "01/06/2020"
output: pdf_document
toc: true
---

# Reference
Zecca, E., Brunelli, C., Centurioni, F., Manzoni, A., Pigni, A., & Caraceni, A. (2017). Fentanyl Sublingual Tablets Versus Subcutaneous Morphine for the Management of Severe Cancer Pain Episodes in Patients Receiving Opioid Treatment: A Double-Blind, Randomized, Noninferiority Trial. *Journal of Clinical Oncology*, 35(7), 759–765. https://doi.org/10.1200/JCO.2016.69.9504

```{r, include=FALSE}
# install.packages("biostatUZH", repos="http://R-Forge.R-project.org")
library(biostatUZH)
library(car)
library(readxl)
library(foreign)
library(readxl)
library(openxlsx)
library(ggplot2)
library(effects)
```

# Notes from reading methods section
* Dependent variable: Pain intensity (PI) (average across follow-ups, AVP_30)
* Independent variable: treatment group
  * Fentanyl sublingual tablets FST (n=57, 1 from 58 was excluded)
  * Subcutaneous morphine SCM (n=56)
* Covariate: baseline PI
* Design: 1-way ANCOVA with group as IV and baseline PI as covariate

# Reading data
Data is loaded, reshaped if necessary, and factors are specified.
```{r}
PATH = file.path(path.expand("~"), "Data", "ancova") # ancova project folder
data = read_excel(file.path(PATH, "dataPrimaryStudies", "03-Zecca", "03-Zecca.xls"))
data = as.data.frame(data)
data$tratt = as.factor(data$tratt)
```

# Descriptives
## Dependent variable
Number of samples and mean (SD) in levels of the independent variables.
```{r}
tab.dv = array(NA, dim=c(2,3))
tab.dv[,1] = levels(data$tratt)
tab.dv[,2] = summary(data$tratt)
tab.dv[,3] = tapply(data$pain_mea_30, data$tratt, 
                  function (x) sprintf("%0.2f (%0.2f)", mean(x), sd(x)))
colnames(tab.dv) = c("group", "n", "mean (SD)")
print(tab.dv)
```

```{r, fig.height=2, fig.width=3.3}
ggplot(data, aes(y=pain_mea_30, x=tratt, color=tratt)) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.15, height = 0)) +
  theme_minimal() +
  theme(axis.text.x = element_blank(), legend.title = element_blank()) +
  xlab("group") + ylab("mean PI across follow-ups ") + ggtitle("Dependant variable")
```

## Covariate(s)
```{r}
tab.cv = array(NA, dim=c(2,3))
tab.cv[,1] = levels(data$tratt)
tab.cv[,2] = summary(data$tratt)
tab.cv[,3] = tapply(data$pain_mea_bas, data$tratt, 
                  function (x) sprintf("%0.2f (%0.2f)", mean(x), sd(x)))
colnames(tab.cv) = c("group", "n", "mean (SD)")
print(tab.cv)
```

```{r fig.height=2, fig.width=3.5}
ggplot(data, aes(y=pain_mea_bas, x=tratt, color=tratt)) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.15, height = 0)) +
  theme_minimal() +
  theme(axis.text.x = element_blank(), legend.title = element_blank()) +
  xlab("group") + ylab("PI at baseline") + ggtitle("Covarate")
```

# Main analysis ANCOVA
```{r, include=FALSE}
stats.orig.IV = data.frame(
  Fvalue =  NA,
  df1 = NA,
  df2 = NA,
  pvalue = NA,
  MD = NA,
  lowerCI = NA,
  upperCI = NA)

stats.orig.CV = stats.orig.IV

# enter all results from primary study here
stats.orig.IV$MD = -0.49
stats.orig.IV$lowerCI = -1.10
stats.orig.IV$upperCI = 0.09
```

```{r}
# Orthogonal contrasts
contrasts(data$tratt) = contr.helmert(2) 

fit.ancova = aov(pain_mea_30 ~ pain_mea_bas + tratt, data = data)
result = Anova(fit.ancova, type=3) # Type III
print(result)
```

The paper used ANCOVA to to a non-inferiority analysis. We calculate ANCOVA adjusted means and 95%-CI.
```{r}
n = summary(data$tratt)
means.adj = effect("tratt", fit.ancova, se=T)
MD = means.adj$fit[1] - means.adj$fit[2] 
# variance assumed unequal
SE = sqrt(means.adj$se[1]^2 + means.adj$se[2]^2)
#ci95 = c(MD - 1.96*SE, MD + 1.96*SE)
# SE based on t-distribution because n1 and n2 < 60
myt = qt(.975, df= n[1] + n[2] - 2)
ci95 = c(MD - myt*SE, MD + myt*SE)

sprintf("%.2f (%.2f, %.2f)", MD, ci95[1], ci95[2])
```

```{r, include=FALSE}
stats.rep.IV = data.frame(Fvalue = sprintf("%.2f",result$`F value`[3]),
                          df1 = result$Df[3],
                          df2 = result$Df[4],
                          pvalue = formatPval(result$`Pr(>F)`[3]),
                          MD = sprintf("%.2f", MD),
                          lowerCI = sprintf("%.2f", ci95[1]),
                          upperCI = sprintf("%.2f", ci95[2])
)

stats.rep.CV = data.frame(Fvalue = sprintf("%.2f",result$`F value`[2]),
                          df1 = result$Df[2],
                          df2 = result$Df[4],
                          pvalue = formatPval(result$`Pr(>F)`[2]),
                          MD = NA,
                          lowerCI = NA,
                          upperCI = NA
)
```

# Comparing ANCOVA in original study with reanalysis
## Independent variable
```{r}
tab.IV = rbind(stats.orig.IV, stats.rep.IV)
rownames(tab.IV) = c("original Study", "reanalysis")
print(t(tab.IV))
```

## Covariate
```{r}
tab.CV = rbind(stats.orig.CV, stats.rep.CV)
rownames(tab.CV) = c("original Study", "reanalysis")
print(t(tab.CV))
```

# Assumptions
## 1.  Homogeneity of variance
* ANOVA/ANCOVA is fairly robust in terms of the error rate when sample sizes are equal.
* When groups with larger sample sizes have larger variances than the groups with smaller sample sizes, the resulting F-ratio tends to be conservative. That is, it's more likely to produce a non-significant result when a genuine difference does exist in the population.
* Conversely, when the groups with larger sample sizes have smaller variances than the groups with smaller samples sizes, the resulting F-ratio tends to be liberal and can inflate the false positive rate.
```{r}
tapply(data$pain_mea_30, data$tratt, sd)
leveneTest(pain_mea_30 ~ tratt, data = data)
```


## 2.  Independence between covariate and IV
When the covariate and the experimental effect (independent variable) are not independent the treatment effect is obscured, spurious treatment effects can arise and the interpretation of the ANCOVA is seriously compromised.

We test whether our groups differ on the CV. If the groups do not significantly differ then is appropriate to use the covariate.
```{r}
fit.cv = aov(pain_mea_bas ~ tratt, data = data) 
Anova(fit.cv, type=3)
```


## 3. Homogeneity of regression slopes
```{r}
fit.hrs = aov(pain_mea_30 ~ pain_mea_bas*tratt, data = data)
Anova(fit.hrs, type=3)
```

```{r fig.height=2, fig.width=3.5}
ggplot(data, aes(y=pain_mea_30, x=pain_mea_bas, color=tratt, shape=tratt)) +
  geom_point() +
  geom_smooth(formula = y ~ x,method=lm, se=FALSE, fullrange=TRUE) +
  theme_minimal() +
  theme(legend.title = element_blank()) +
  xlab("pain at baseline") + ylab("pain follow-up")
```

# Notes

* Non-inferiority intervention study, ANCOVA is used to calculate adjusted means for the two groups (corrected for baseline pain)
* Main finding partly reproduced, lower bound of the 95%-CI was wrong, probably a typo.
* Our analysis was consistent with SAS output from authors
* All assumptions met

Data was analyzed according to recommendations by Field, Miles, & Field (2012).