---
title: "Reanalysis of 11-Carlisle"
author: "Simon Schwab^[University of Zurich, simon.schwab@uzh.ch]"
date: "26/08/2020"
output: pdf_document
toc: true
---

# Reference
Carlisle, A., Selwood, L., Hinds, L. A., Saunders, N., Habgood, M., Mardon, K., & Weisbecker, V. (2017). Testing hypotheses of developmental constraints on mammalian brain partition evolution, using marsupials. Scientific Reports, 7(1), 4241. https://doi.org/10.1038/s41598-017-02726-9

```{r, include=FALSE}
# install.packages("biostatUZH", repos="http://R-Forge.R-project.org")
library(biostatUZH)
library(car)
library(readxl)
library(foreign)
library(readxl)
library(openxlsx)
library(ggplot2)
library(effects)
```

# Notes from reading methods section

* Dependent variable: brain partition size of olfactory bulb
* Independent variable: species (3 levels)
  * Macropus eugenii
  * Trichosurus vulpecula
  * Monodelphis domestica
* Covariate: whole brain volume
* Design: 1-way ANCOVA with species as IV and whole brain volume as covariate

# Reading data
Data is loaded, reshaped if necessary, and factors are specified.
```{r}
PATH = file.path(path.expand("~"), "Data", "ancova") # ANCOVA project folder
data = read.csv(file.path(PATH, "dataPrimaryStudies", "11-Carlisle", "11-Carlisle-original.txt"),
                sep = "", header = TRUE)

data$Species.name = NA
data$Species.name[data$Species == 1] = "Macropus eugenii"
data$Species.name[data$Species == 2] = "Trichosurus vulpecula"
data$Species.name[data$Species == 3] = "Monodelphis domestica"

data$Species       = as.factor(data$Species)
data$Species.name = as.factor(data$Species.name)

# first reported analyses are only with juveniles
data = subset(data, subset = Juvenile.Adult == "juvenile")
```

# Descriptives
## Dependant variable
Number of samples and mean (SD) in levels of the independent variables.
```{r}
tab.dv = array(NA, dim=c(nlevels(data$Species.name),3))
tab.dv[,1] = levels(data$Species.name)
tab.dv[,2] = summary(data$Species.name)
tab.dv[,3] = tapply(data$Olfactory_Bulb, data$Species.name, 
                  function (x) sprintf("%0.2f (%0.2f)", mean(x), sd(x)))
colnames(tab.dv) = c("group", "n", "mean (SD)")
print(tab.dv)
```

```{r, fig.height=2, fig.width=3.3}
ggplot(data, aes(y=log(Olfactory_Bulb), x=Species.name, color=Species.name)) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.15, height = 0)) +
  theme_minimal() +
  theme(axis.text.x = element_blank(), legend.title = element_blank()) +
  xlab("species") + ylab("olfactory bulb volume (log)") + ggtitle("Dependant variable")
```

## Covariate(s)
```{r}
tab.cv = array(NA, dim=c(nlevels(data$Species.name),3))
tab.cv[,1] = levels(data$Species.name)
tab.cv[,2] = summary(data$Species.name)
tab.cv[,3] = tapply(data$Whole_Brain_Volume, data$Species.name, 
                  function (x) sprintf("%0.2f (%0.2f)", mean(x), sd(x)))
colnames(tab.cv) = c("group", "n", "mean (SD)")
print(tab.cv)
```

```{r fig.height=2, fig.width=3.5}
ggplot(data, aes(y=log(Whole_Brain_Volume), x=Species.name, color=Species.name)) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.15, height = 0)) +
  theme_minimal() +
  theme(axis.text.x = element_blank(), legend.title = element_blank()) +
  xlab("group") + ylab("Whole brain volume (log)") + ggtitle("Covarate")
```

# Main analysis ANCOVA
```{r, include=FALSE}
stats.orig.IV = data.frame(
  Fvalue =  NA,
  df1 = NA,
  df2 = NA,
  pvalue = NA,
  MD = NA,
  lowerCI = NA,
  upperCI = NA)

stats.orig.CV = stats.orig.IV

# enter all results from primary study here
stats.orig.IV$pvalue = "< 0.0001"
```

```{r}
# Orthogonal contrasts
contrasts(data$Species.name) = contr.helmert(3) 

# The author did not run a classic ANCOVA with IV + COV, but instead an interaction
# COV*IV similar to testing the homogeneity of slope assumption
fit.ancova = aov(log(Olfactory_Bulb) ~ log(Whole_Brain_Volume-Olfactory_Bulb)*Species, data = data)
result = Anova(fit.ancova, type=3) # Type III
print(result)
```


```{r, include=FALSE}
stats.rep.IV = data.frame(Fvalue = sprintf("%.2f",result$`F value`[4]),
                          df1 = result$Df[4],
                          df2 = result$Df[5],
                          pvalue = formatPval(result$`Pr(>F)`[4]),
                          MD = NA,
                          lowerCI = NA,
                          upperCI = NA
)

stats.rep.CV = data.frame(Fvalue = sprintf("%.2f",result$`F value`[2]),
                          df1 = result$Df[2],
                          df2 = result$Df[4],
                          pvalue = formatPval(result$`Pr(>F)`[2]),
                          MD = NA,
                          lowerCI = NA,
                          upperCI = NA
)
```

# Comparing ANCOVA in original study with reanalysis
## Independent variable
```{r}
tab.IV = rbind(stats.orig.IV, stats.rep.IV)
rownames(tab.IV) = c("original Study", "reanalysis")
print(t(tab.IV))
```

## Covariate
```{r}
tab.CV = rbind(stats.orig.CV, stats.rep.CV)
rownames(tab.CV) = c("original Study", "reanalysis")
print(t(tab.CV))
```

# Assumptions
## 1.  Homogeneity of variance
* ANOVA/ANCOVA is fairly robust in terms of the error rate when sample sizes are equal.
* When groups with larger sample sizes have larger variances than the groups with smaller sample sizes, the resulting F-ratio tends to be conservative. That is, it's more likely to produce a non-significant result when a genuine difference does exist in the population.
* Conversely, when the groups with larger sample sizes have smaller variances than the groups with smaller samples sizes, the resulting F-ratio tends to be liberal and can inflate the false positive rate.
```{r}
tapply(log(data$Olfactory_Bulb), data$Species.name, sd)
leveneTest(log(Olfactory_Bulb) ~ Species.name, data = data)
```


## 2.  Independence between covariate and IV
* When the covariate and the experimental effect (independent variable) are not independent the treatment effect is obscured, spurious treatment effects can arise and the interpretation of the ANCOVA is seriously compromised.

* We test whether our groups differ on the CV. If the groups do not significantly differ then is appropriate to use the covariate.
```{r}
fit.cv = aov(log(Whole_Brain_Volume-Olfactory_Bulb) ~ Species.name, data = data) 
Anova(fit.cv, type=3)
```


## 3. Homogeneity of regression slopes
* We test the interaction between the IV and the CV
```{r}
fit.hrs = aov(log(Olfactory_Bulb) ~ log(Whole_Brain_Volume-Olfactory_Bulb)*Species.name, data = data)
Anova(fit.hrs, type=3)
```

```{r fig.height=2, fig.width=3.5}
ggplot(data, aes(y=log(Olfactory_Bulb), x=log(Whole_Brain_Volume-Olfactory_Bulb), color=Species.name, shape=Species.name)) +
  geom_point() +
  geom_smooth(formula = y ~ x,method=lm, se=FALSE, fullrange=TRUE) +
  theme_minimal() +
  theme(legend.title = element_blank()) +
  xlab("whole brain volume") + ylab("olfactory bulb")
```

# Additional analyses
```{r fig.height=2, fig.width=2}
# Reproducing Figure 4
ggplot(data, aes(x = log(Whole_Brain_Volume), y = log(Olfactory_Bulb))) +
  geom_point() + theme_minimal()
# Reproducing Table 1 
fit = lm(log(Olfactory_Bulb)~log(Whole_Brain_Volume-Olfactory_Bulb), data=data)
summary(fit)

fit = lm(log(Olfactory_Bulb)~log(Whole_Brain_Volume-Olfactory_Bulb), 
         data=subset(data, subset = Species == 1))
summary(fit)
```

# Notes

* First main result reproduced
* Assumptions hold. Homogeneity of regression slopes was part of the research questions (Table 2).
* ANCOVA used with interactions term for species and whole brain volume, not to (globally) regress covariate
* Log transform of IV and CV was successful
* Helpful R code was supplied
* Model was clearly specified in the methods section
* Explain ANCOVA in the framework of regression (slope, intercept)
* Table 2 is confusing as p-values were associated with the DV instead of the IV
* Lots of models tested, no correction for multiplicity (7 ANCOVAs were run across different brain regions)

Data was analyzed according to recommendations by Field, Miles, & Field (2012).