---
title: "Reanalysis of 17-DePascalis"
author:
- Audrey Yeo^[University of Zurich, audrey.yeo@uzh.ch]
- Simon Schwab^[University of Zurich, simon.schwab@uzh.ch]
date: "31/08/2020"
output: pdf_document
toc: true
---

# Reference
De Pascalis, V., Fracasso, F., & Corr, P. J. (2017). Personality and Augmenting/Reducing (A/R) in auditory event-related potentials (ERPs) during emotional visual stimulation. Scientific Reports, 7, 41588. https://doi.org/10.1038/srep41588

```{r, include = FALSE}
# install.packages("biostatUZH", repos="http://R-Forge.R-project.org")
library(biostatUZH)
library(car)
library(readxl)
library(foreign)
library(readxl)
library(ggplot2)
library(cowplot)
```

# Notes from reading methods section
* Dependant variable: "N1/P2 complex" (n = 39)
* Independent variables:
  * Emotion, 3 levels (negative, neutral, positive pictures)
  * Auditory intensity, 5 levels (59, 70, 79, 88, 96 dB SPL)
  * Recording Site, 3 levels (frontal Fz, central Cz, parietal Pz)
* Covariate: BIS (Behavioural Inhibition System) with RST-PQ questionnaire
* Design: 3-way 3x5x3 ANCOVA (all factors within)

# Reading data
Data is loaded, reshaped if necessary, and factors are specified.
```{r message=FALSE}
PATH = file.path(path.expand("~"), "Data", "ancova") # ancova project folder
raw = read_excel(file.path(PATH, "dataPrimaryStudies", "17-DePascalis", "BIS_cov_N1-P2peak_amplitude_r.xlsx"), range = "B5:AW44", col_names = TRUE)
raw = as.data.frame(raw)
# correcting typo
colnames(raw)[1] = "Subject"
```

## Reshaping data into long
```{r}
N = 39
data = data.frame(subject = as.factor(rep(raw$Subject, 3*5*3)))
data$amplitude = c(as.matrix(raw[,2:46]))
data$emotion   = as.factor(c(rep("negative", 5*3*N), rep("neutral", 5*3*N), rep("positive", 5*3*N)))
data$intensity = as.factor(rep(c(rep("level 1", 3*N), rep("level 2", 3*N), rep("level 3", 3*N), rep("level 4", 3*N), rep("level 5", 3*N)), 3))
data$site      = as.factor(rep(c(rep("location 1", N), rep("location 2", N), rep("location 3", N)), 3*5))
data$bis       = rep(raw$RST_BIS, 3*5*3)
summary(data)
```
Running an initial ANCOVA

# Descriptives
## Dependent variable
We show descriptives plots and no tables as the factorial designs is rather complex (3x5x3).

### Main effects
Averaging subjects across the other two levels
```{r fig.height=2, fig.width=11}
# 
d.1 = aggregate(amplitude ~ bis + emotion*subject, data = data, FUN = mean)
d.2 = aggregate(amplitude ~ bis + intensity*subject, data = data, FUN = mean)
d.3 = aggregate(amplitude ~ bis + site*subject, data = data, FUN = mean)

p1 = ggplot(d.1, aes(y=amplitude, x=emotion, color=emotion)) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.15, height = 0)) +
  theme_minimal()

p2 = ggplot(d.2, aes(y=amplitude, x=intensity, color=intensity)) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.15, height = 0)) +
  theme_minimal()

p3 = ggplot(d.3, aes(y=amplitude, x=site, color=site)) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.15, height = 0)) +
  theme_minimal()

plot_grid(p1, p2, p3, nrow = 1, ncol = 3)
```
### Interaction effects (two-way)
Averaging subjects across along the third factor to see the two-way interactions. Plots show means (SD as error bar)
```{r fig.height=2, fig.width=11}
d.A = aggregate(amplitude ~ emotion*intensity, data = data, FUN = mean)
d.B = aggregate(amplitude ~ emotion*site, data = data, FUN = mean)
d.C = aggregate(amplitude ~ intensity*site, data = data, FUN = mean)

# get SD
d.A$sd = aggregate(amplitude ~ emotion*intensity, data = data, FUN = sd)[,3]
d.B$sd = aggregate(amplitude ~ emotion*site, data = data, FUN = sd)[,3]
d.C$sd = aggregate(amplitude ~ intensity*site, data = data, FUN = sd)[,3]

pA = ggplot(d.A, aes(y=amplitude, x=intensity, group=emotion, color=emotion)) +
  geom_errorbar(aes(ymin=amplitude-sd, ymax=amplitude+sd), width=.1) +
  geom_line() + geom_point() +
  theme_minimal()

pB = ggplot(d.B, aes(y=amplitude, x=site, group=emotion, color=emotion)) +
  geom_errorbar(aes(ymin=amplitude-sd, ymax=amplitude+sd), width=.1) +
  geom_line() + geom_point() +
  theme_minimal()

pC = ggplot(d.C, aes(y=amplitude, x=intensity, group=site, color=site)) +
  geom_errorbar(aes(ymin=amplitude-sd, ymax=amplitude+sd), width=.1) +
  geom_line() + geom_point() +
  theme_minimal()

plot_grid(pA, pB, pC, nrow = 1, ncol = 3)
```

## Covariate(s)
Descriptives, see Table 1
```{r}
sprintf("BIS: %0.1f (%0.1f)", mean(data$bis[1:N]), sd(data$bis[1:N]))
```

# Main analysis ANCOVA
```{r, include=FALSE}
stats.orig.IV = data.frame(
  Fvalue =  NA,
  df1 = NA,
  df2 = NA,
  pvalue = NA,
  MD = NA,
  lowerCI = NA,
  upperCI = NA)

stats.orig.CV.bis = data.frame(
  Fvalue =  6.82,
  df1 = 1,
  df2 = 37,
  pvalue = 0.013,
  MD = NA,
  lowerCI = NA,
  upperCI = NA)
```


```{r}
# Orthogonal contrasts
# contrasts(data$group) = contr.helmert(2)

fit.ancova = aov(amplitude ~ bis + emotion*intensity*site + Error(subject/(emotion*intensity*site)), data = data)
result = summary(fit.ancova) # Type I
#result = Anova(fit.ancova, type=3) # Type III
print(result)
```

```{r, include=FALSE}
stats.rep.IV = data.frame(Fvalue = NA,
                          df1 = NA,
                          df2 = NA,
                          pvalue = NA,
                          MD = NA,
                          lowerCI = NA,
                          upperCI = NA
)

stats.rep.CV.bis = data.frame(Fvalue = sprintf("%.2f",result$`Error: subject`[[1]]$`F value`[1]),
                          df1 = result$`Error: subject`[[1]]$Df[1],
                          df2 = result$`Error: subject`[[1]]$Df[2],
                          pvalue = formatPval(result$`Error: subject`[[1]]$`Pr(>F)`[1]),
                          MD = NA,
                          lowerCI = NA,
                          upperCI = NA
)
```

# Comparing ANCOVA in original study with reanalysis
## Independant variable
```{r}
tab.IV = rbind(stats.orig.IV, stats.rep.IV)
rownames(tab.IV) = c("original Study", "reanalysis")
print(t(tab.IV))
```

## Covariate
```{r}
tab.CV = rbind(stats.orig.CV.bis, stats.rep.CV.bis)
rownames(tab.CV) = c("original Study", "reanalysis")
print(t(tab.CV))
```

# Assumptions
## 1.  Homogeneity of variance
* ANOVA/ANCOVA is fairly robust in terms of the error rate when sample sizes are equal.
* When groups with larger sample sizes have larger variances than the groups with smaller sample sizes, the resulting F-ratio tends to be conservative. That is, it's more likely to produce a non-significant result when a genuine difference does exist in the population.
* Conversely, when the groups with larger sample sizes have smaller variances than the groups with smaller samples sizes, the resulting F-ratio tends to be liberal and can inflate the false positive rate.
```{r}
tmp = tapply(d.1$amplitude, d.1$emotion, sd)
tab = sprintf("%.2f", tmp)
names(tab) = names(tmp)
stats = leveneTest(amplitude ~ emotion, data = d.1)
tab[length(tab)+1] = formatPval(stats$`Pr(>F)`[1])
names(tab)[length(tab)] = "Levene's p-value"
print(tab)

tmp = tapply(d.2$amplitude, d.2$intensity, sd)
tab = sprintf("%.2f", tmp)
names(tab) = names(tmp)
stats = leveneTest(amplitude ~ intensity, data = d.2)
tab[length(tab)+1] = formatPval(stats$`Pr(>F)`[1])
names(tab)[length(tab)] = "Levene's p-value"
print(tab)

tmp = tapply(d.3$amplitude, d.3$site, sd)
tab = sprintf("%.2f", tmp)
names(tab) = names(tmp)
stats = leveneTest(amplitude ~ site, data = d.3)
tab[length(tab)+1] = formatPval(stats$`Pr(>F)`[1])
names(tab)[length(tab)] = "Levene's p-value"
print(tab)

```


## 2.  Independence between covariate and IV
When the covariate and the experimental effect (independent variable) are not independent the treatment effect is obscured, spurious treatment effects can arise and the interpretation of the ANCOVA is seriously compromised.

We test whether our groups differ on the CV. If the groups do not significantly differ then is appropriate to use the covariate.
```{r}
fit.cv = aov(bis ~ emotion, data = d.1) 
summary(fit.cv)
```

Not applicable for a within design as the covariate does not change across levels. So this assumption does not need testing.

## 3. Homogeneity of regression slopes
* We test the interaction between the IV and the CV
```{r}
fit.hrs = aov(amplitude ~ bis*emotion + Error(subject/emotion), data = d.1)
summary(fit.hrs)
```

```{r}
fit.hrs = aov(amplitude ~ bis*intensity + Error(subject/intensity), data = d.2)
summary(fit.hrs)
```

```{r}
fit.hrs = aov(amplitude ~ bis*site + Error(subject/site), data = d.3)
summary(fit.hrs)
```

```{r fig.height=2, fig.width=11}
p1 = ggplot(d.1, aes(y=amplitude, x=bis, color=emotion, shape=emotion)) +
  geom_point() +
  geom_smooth(formula = y ~ x,method=lm, se=FALSE, fullrange=TRUE) +
  theme_minimal() +
  theme(legend.title = element_blank())

p2 = ggplot(d.2, aes(y=amplitude, x=bis, color=intensity, shape=intensity)) +
  geom_point() +
  geom_smooth(formula = y ~ x,method=lm, se=FALSE, fullrange=TRUE) +
  theme_minimal() +
  theme(legend.title = element_blank())

p3 = ggplot(d.3, aes(y=amplitude, x=bis, color=site, shape=site)) +
  geom_point() +
  geom_smooth(formula = y ~ x,method=lm, se=FALSE, fullrange=TRUE) +
  theme_minimal() +
  theme(legend.title = element_blank())

plot_grid(p1, p2, p3, ncol = 3, nrow = 1)
```

# Notes

* Amplitudes were not normal, log transform could have had helped.
* The first reported ANCOVA result was reproduced: effect of the covariate BIS
* BIS was used as continuous covariate which showed a statistically sign. effect. on amplitude. However, authors reported an effect of low vs. high BIS participants on amplitude (categorical), with lower amplitudes in high BIS participants based on Figure 2.
* Not clear how low vs. high BIS was determined (which threshold was used?).
* Only parts of ANCOVA results were reported. There were altogether three main effects and three two-way interactions that were not all reported.
* Generally, there is a mismatch between the ANCOVA model specified and the result reported (BIS treated continuous or categorical?).
* Assumptions of homogeneity of regression slope was not met for intensity and emotion: the relationship between BIS and amplitude differed across levels of intensity and levels of emotion. The authors reported both, so this is fine.
* Assumption of independence between COV and IV cannot be tested in within-only designs.

Data was analyzed according to recommendations by Field, Miles, & Field (2012).
