---
title: "19-Pallauf"
author: "Audrey Yeo^[University of Zurich, audrey.yeo@uzh.ch]"
date: "01/06/2020"
output: pdf_document
toc: true
---

# Reference
Pallauf, K., Chin, D., Günther, I. et al. Resveratrol, lunularin and dihydroresveratrol do not act as caloric restriction mimetics when administered intraperitoneally in mice. *Sci Rep* 9, 4445 (2019). https://doi.org/10.1038/s41598-019-41050-2

```{r, include=FALSE}
# install.packages("biostatUZH", repos="http://R-Forge.R-project.org")
library(biostatUZH)
library(car)
library(readxl)
library(foreign)
library(readxl)
library(ggplot2)
```

# Notes from reading methods section
An analysis of covariances (ANCOVA) was con- ducted for the weight development. For the weekly weight development, the statistical model included the treatment group (CON, RSV, DHR, LUN), the covariate ‘week’ as well as their interaction term as fixed factors. The mouse was regarded as random factor. We excluded the CR mice from the statistical analysis of feed uptake and bodyweight.
 
* Dependant variable: Weight
* Independant variable: Treatment group 
  * CON, 
  * RSV, 
  * DHR, 
  * LUN
* Covariate: Time

# Reading data
Data is loaded, reshaped if necessary, and factors are specified.
```{r}
PATH = file.path(path.expand("~"), "Desktop", "CRS", "ancova") # ancova project folder
data = read.csv(file.path(PATH, "dataPrimaryStudies", "19-Pallauf", "19-Pallauf-1.csv"))
# data = read.csv("~CRS/ancova/dataPrimaryStudies/19-Pallauf/19-Pallauf-1.csv")
names(data)
class(data$Group)
# data$Group <- data$Group %in% c("CON", "CR", "DHR", "LUN", "RES")
data$Group.factor = NA
data$Group.factor[data$Group == "CR"] = "CR"
data$Group.factor[data$Group == "DHR"] = "DHR"
data$Group.factor[data$Group == "CON"] = "CON"
data$Group.factor[data$Group == "LUN"] = "LUN"
data$Group.factor[data$Group == "RES"] = "RES"
sum(is.na(data$Weight)) #33 missing values
data$Group.factor = factor(data$Group.factor,
                           levels = c("CR", "DHR",
                                      "CON", "LUN", "RES"))
```

# Descriptives
## Dependant variable
Number of samples and mean (SD) in levels of the independant variables. We reproduce Table 3 and Figure 2A of the study.
```{r}
idx = c(1, 3, 2, 4) # sorting as in publication
tab.dv = array(NA, dim=c(6,3))
tab.dv[,1] = levels(data$Group)
tab.dv[,2] = summary(data$Group)
tab.dv[,3] = tapply(data$Weight, data$Group, 
                  function (x) sprintf("%0.2f (%0.2f)", mean(x, na.rm = TRUE), sd(x, na.rm = TRUE)))
colnames(tab.dv) = c("group", "n", "mean (SD)")
print(tab.dv)
```

```{r fig.height=2, fig.width=3.5}
ggplot(na.omit(data), aes(y=Weight, x=Group.factor, color=Group.factor)) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.15, height = 0)) +
  theme_minimal() +
  theme(axis.text.x = element_blank(), legend.title = element_blank()) +
  xlab("group") + ylab("Weight") + ggtitle("Dependant variable")
```

## Covariate(s)
```{r}
idx = c(1, 3, 2, 4) # sorting as in publication
tab.cv = array(NA, dim=c(6,3))
tab.cv[,1] = levels(data$Group)
tab.cv[,2] = summary(na.omit(data$Group))
tab.cv[,3] = tapply(data$Weight, data$Group, 
                  function (x) sprintf("%0.1f (%0.1f)", mean(x), sd(x)))
colnames(tab.cv) = c("group", "n", "mean (SD)")
print(tab.cv)
```

```{r fig.height=2, fig.width=3.5}
ggplot(data, aes(y=Weight, x=Week, color=Group.factor)) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.15, height = 0)) +
  theme_minimal() +
  theme(axis.text.x = element_blank(), legend.title = element_blank()) +
  xlab("Week") + ylab("Weight") + ggtitle("Covariate")
```

# Main analysis ANCOVA
```{r, include=FALSE}
stats.orig.IV = data.frame(
  Fvalue =  NA,
  df1 = NA,
  df2 = NA,
  pvalue = "p < 0.001",
  MD = NA,
  lowerCI = NA,
  upperCI = NA)

stats.orig.CV = stats.orig.IV

# enter all results from primary study here
stats.orig.IV$pvalue = "n.s." 
```


```{r}
# Orthogonal contrasts
contrasts(data$Group) = contr.helmert(6) 

fit.ancova = aov(Weight ~ Week + Group, data = data)
# result = summary(fit) # Type I
result = Anova(fit.ancova, type=3) # Type III
print(result)
```

```{r, include=FALSE}
stats.rep.IV = data.frame(Fvalue = sprintf("%.2f",result$`F value`[3]),
                          df1 = result$Df[3],
                          df2 = result$Df[4],
                          pvalue = formatPval(result$`Pr(>F)`[3]),
                          MD = NA,
                          lowerCI = NA,
                          upperCI = NA
)

stats.rep.CV = data.frame(Fvalue = sprintf("%.2f",result$`F value`[2]),
                          df1 = result$Df[2],
                          df2 = result$Df[4],
                          pvalue = formatPval(result$`Pr(>F)`[2]),
                          MD = NA,
                          lowerCI = NA,
                          upperCI = NA
)
```

# Comparing ANCOVA in original study with reanalysis
## Independant variable
```{r}
tab.IV = rbind(stats.orig.IV, stats.rep.IV)
rownames(tab.IV) = c("original Study", "reanalysis")
print(t(tab.IV))
```

## Covariate
```{r}
tab.CV = rbind(stats.orig.CV, stats.rep.CV)
rownames(tab.CV) = c("original Study", "reanalysis")
print(t(tab.CV))
```

# Assumptions
## 1.  Homogeneity of variance
* ANOVA/ANCOVA is fairly robust in terms of the error rate when sample sizes are equal.
* When groups with larger sample sizes have larger variances than the groups with smaller sample sizes, the resulting F-ratio tends to be conservative. That is, it's more likely to produce a non-significant result when a genuine difference does exist in the population.
* Conversely, when the groups with larger sample sizes have smaller variances than the groups with smaller samples sizes, the resulting F-ratio tends to be liberal and can inflate the false positive rate.
```{r}
tapply(data$Weight, data$Group, sd)
leveneTest(Weight ~ Group, data = data)
```


## 2.  Independence between covariate and IV
When the covariate and the experimental effect (independent variable) are not independent the treatment effect is obscured, spurious treatment effects can arise and the interpretation of the ANCOVA is seriously compromised.

We test whether our groups differ on the CV. If the groups do not significantly differ then is appropriate to use the covariate.
```{r}
fit.cv = aov(Weight ~ Group, data = data) 
summary(fit.cv)
```


## 3. Homogeneity of regression slopes
```{r}
fit.hrs = aov(Weight ~ Week*Group, data = data)
Anova(fit.hrs, type=3)
```

```{r fig.height=2, fig.width=3.5}
ggplot(data, aes(y=Weight, x=Week, color=Group, shape=Group)) +
  geom_point() +
  geom_smooth(formula = y ~ x,method=lm, se=FALSE, fullrange=TRUE) +
  theme_minimal() +
  theme(legend.title = element_blank()) +
  xlab("Week") + ylab("Weight")
```



# Notes

* The first reported ANCOVA which was n.s. was reproduced
* p-value was not reported
* Authors also report age unadjusted analyses and report these as ANCOVAs instead of ANOVAs (p. 357) which was confusing
* Assumptions were met, except homogeneity of regression slopes showing a different relationship of outcome and covariate in the LRRK2 premanifest group
* Many analyses, they corrected for multiplicity

Data was analyzed according to recommendations by Field, Miles, & Field (2012).