---
title: "Reanalysis of 20-Castagneyrol"
author: "Simon Schwab^[University of Zurich, simon.schwab@uzh.ch]"
date: "18/03/2021"
output: pdf_document
toc: true
---

# Reference
Castagneyrol, B., Moreira, X., & Jactel, H. (2018). Drought and plant neighbourhood interactively determine herbivore consumption and performance. Scientific Reports, 8(1), 5930. https://doi.org/10.1038/s41598-018-24299-x

```{r, include=FALSE}
# install.packages("biostatUZH", repos="http://R-Forge.R-project.org")
library(biostatUZH)
library(car)
library(readxl)
library(foreign)
library(readxl)
library(ggplot2)
library(doBy)
library(cowplot)
```

# Notes from reading methods section
* Dependant variable: Relative Consumption Rate (RCR) (log transformed)
* Independent variable: 
  * Irrigation (non-irrigated vs irrigated)
  * Composition (B, BG, BP, BQP)
* Covariate: initial weight, w_i
* Design: 2-way 2x4 ANCOVA irrigation x composition (both between) as IV and initial weight as covariate
* N = 96

# Reading data
Data is loaded, reshaped if necessary, and factors are specified.
```{r}
PATH = file.path(path.expand("~"), "Data", "ancova") # ancova project folder

d.leaf.lab = read.csv(file.path(PATH, "dataPrimaryStudies", "20-Castagneyrol", "d_leaf_lab.csv"), header = TRUE, sep = "\t")
d.leaf.lab$Composition = factor(d.leaf.lab$Composition, levels = c('B', 'BQ', 'BP', 'BQP'))
dc = read.csv(file.path(PATH, "dataPrimaryStudies", "20-Castagneyrol", "dc.csv"), header = TRUE, sep = "\t")
N = 96
```
## Preprocessing
Authors preprocessed data to create the outcome variable $Consumption$, and combed the two datasets (code by original author code)
```{r}
## one removed outlier for lava weight
d.leaf.lab$WL.t1[d.leaf.lab$WL.t1 > 0.5] <- NA

d.leaf.lab$Treatment <- as.factor(paste(d.leaf.lab$Irrigation, d.leaf.lab$Composition, sep= '_'))
a <- unlist(lapply(split(d.leaf.lab, d.leaf.lab$Treatment), function(x){coef(lm(WL.t1 ~ SL.t1, x))[[2]]})) # extract slope
b <- unlist(lapply(split(d.leaf.lab, d.leaf.lab$Treatment), function(x){coef(lm(WL.t1 ~ SL.t1, x))[[1]]})) # extract intercept

dl <- summaryBy(d.SL ~ Box.ID + Treatment + Irrigation + Composition, data = d.leaf.lab, FUN = sum, keep.names = T)

dl$a <- NA
dl$b <- NA

for(i in 1:length(levels(d.leaf.lab$Treatment))){
  dl[dl$Treatment == levels(d.leaf.lab$Treatment)[i], ]$a <- a[i]
  dl[dl$Treatment == levels(d.leaf.lab$Treatment)[i], ]$b <- b[i]
}

dl$Consumption <- with(dl, b + d.SL*a)
dl$Consumption <- dl$Consumption * 1000

dl <- dl[order(dl$Box.ID),]
dc <- dc[order(dc$Box.ID),]
dc$RGR <- with(dc, ((WCf - WCi)/WCi)/8)
# dotchart(sort(dc$RGR))

dc$RGR[dc$RGR > 1] <- NA

dcl <- data.frame(
  Box.ID = dl$Box.ID,
  Composition = dl$Composition,
  Irrigation = dl$Irrigation,
  Consumption = dl$Consumption / 8,
  WCi = dc$WCi,
  WCf = dc$WCf,
  RGR = dc$RGR,
  Growth = (dc$WCf - dc$WCi),
  Frass = dc$Frass / 8
)
rm(d.leaf.lab, dl, dc)
```
 
```{r}
dcl$Irrigation = as.factor(dcl$Irrigation)
summary(dcl)
```


# Descriptives
## Dependant variable

### Main effects
```{r fig.height=2, fig.width=7}
p1 = ggplot(dcl, aes(y=Consumption, x=Composition, color=Composition)) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.15, height = 0)) +
  theme_minimal() +
  theme(axis.text.x = element_blank(), legend.title = element_blank())

p2 = ggplot(dcl, aes(y=Consumption, x=Irrigation, color=Irrigation)) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.15, height = 0)) +
  theme_minimal() +
  theme(axis.text.x = element_blank(), legend.title = element_blank())

plot_grid(p1, p2, nrow=1, ncol=2)
```

### Interaction effects (two-way)
Mean and SE along the 4x2 factorial design, see Figure 1A.
```{r fig.height=2, fig.width=3.5}
d = aggregate(Consumption ~ Composition*Irrigation, data = dcl, FUN = mean)
d$sd = aggregate(Consumption ~ Composition*Irrigation, data = dcl, FUN = sd)[,3]
d$se = d$sd/sqrt(N/4/2)

ggplot(d, aes(y=Consumption, x=Composition, group=Irrigation, color=Irrigation)) +
  geom_errorbar(aes(ymin=Consumption-se, ymax=Consumption+se), width=.1) +
  geom_line() + geom_point() +
  theme_minimal()
```

## Covariate(s)

```{r fig.height=2, fig.width=7}
p1 = ggplot(dcl, aes(y=WCi, x=Composition, color=Composition)) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.15, height = 0)) +
  theme_minimal() +
  theme(axis.text.x = element_blank(), legend.title = element_blank())

p2 = ggplot(dcl, aes(y=WCi, x=Irrigation, color=Irrigation)) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.15, height = 0)) +
  theme_minimal() +
  theme(axis.text.x = element_blank(), legend.title = element_blank())

plot_grid(p1, p2, nrow=1, ncol=2)
```

### Exclusion of outlier
Largest value in the covariate WCi was excluded, but not declared in the paper.
```{r}
dcl.orig = dcl
dcl = subset(dcl.orig, subset = WCi < 0.07)
```


# Main analysis ANCOVA
```{r, include=FALSE}
stats.orig.IV.irrigation = data.frame(
  Fvalue =  25.55,
  df1 = 1,
  df2 = 86,
  pvalue = "< 0.001",
  MD = NA,
  lowerCI = NA,
  upperCI = NA)

stats.orig.IV.composition = data.frame(
  Fvalue =  0.83,
  df1 = 3,
  df2 = 86,
  pvalue = 0.480,
  MD = NA,
  lowerCI = NA,
  upperCI = NA)

stats.orig.IV.interaction = data.frame(
  Fvalue =  8.66,
  df1 = 3,
  df2 = 86,
  pvalue = "< 0.001",
  MD = NA,
  lowerCI = NA,
  upperCI = NA)

stats.orig.CV = data.frame(
  Fvalue =  40.77,
  df1 = 1,
  df2 = 86,
  pvalue = "< 0.001",
  MD = NA,
  lowerCI = NA,
  upperCI = NA)
```


### ANCOVA
```{r}
# Orthogonal contrasts
contrasts(dcl$Composition) = contr.helmert(4)
contrasts(dcl$Irrigation)  = contr.helmert(2)

fit.ancova = aov(log(Consumption) ~ WCi + Irrigation * Composition, data = dcl)
# result.ancova = summary(fit.ancova) # Type I
result.ancova = Anova(fit.ancova, type=3) # Type III
print(result.ancova)
```

### Regression
In the paper, a regression with $lm()$ was used which corresponds to an ANCOVA type I SS. The interaction gives the same result, but main effects and CV are different in type III SS.
```{r}
fit.lm = lm(log(Consumption) ~ WCi + Irrigation * Composition, data = dcl)
result.lm = anova(fit.lm) # Type I SS
# result.lm = Anova(fit.lm, type=3) # Type III SS
print(result.lm)
```

```{r, include=FALSE}
# anova Type 3 SS
stats.rep.IV.irrigation = data.frame(Fvalue = sprintf("%.2f", result.ancova$`F value`[3]),
                          df1 = result.ancova$Df[3],
                          df2 = result.ancova$Df[6],
                          pvalue = formatPval(result.ancova$`Pr(>F)`[3]),
                          MD = NA,
                          lowerCI = NA,
                          upperCI = NA
)

stats.rep.IV.composition = data.frame(Fvalue = sprintf("%.2f", result.ancova$`F value`[4]),
                          df1 = result.ancova$Df[4],
                          df2 = result.ancova$Df[6],
                          pvalue = formatPval(result.ancova$`Pr(>F)`[4]),
                          MD = NA,
                          lowerCI = NA,
                          upperCI = NA
)

stats.rep.IV.interaction = data.frame(Fvalue = sprintf("%.2f", result.ancova$`F value`[5]),
                          df1 = result.ancova$Df[5],
                          df2 = result.ancova$Df[6],
                          pvalue = formatPval(result.ancova$`Pr(>F)`[5]),
                          MD = NA,
                          lowerCI = NA,
                          upperCI = NA
)

stats.rep.CV = data.frame(Fvalue = sprintf("%.2f", result.ancova$`F value`[2]),
                          df1 = result.ancova$Df[2],
                          df2 = result.ancova$Df[6],
                          pvalue = formatPval(result.ancova$`Pr(>F)`[2]),
                          MD = NA,
                          lowerCI = NA,
                          upperCI = NA
)

# lm anova type 1
stats.rep.IV.irrigation.lm = data.frame(Fvalue = sprintf("%.2f", result.lm$`F value`[2]),
                          df1 = result.lm$Df[2],
                          df2 = result.lm$Df[5],
                          pvalue = formatPval(result.lm$`Pr(>F)`[2]),
                          MD = NA,
                          lowerCI = NA,
                          upperCI = NA
)

stats.rep.IV.composition.lm = data.frame(Fvalue = sprintf("%.2f", result.lm$`F value`[3]),
                          df1 = result.lm$Df[3],
                          df2 = result.lm$Df[5],
                          pvalue = formatPval(result.lm$`Pr(>F)`[3]),
                          MD = NA,
                          lowerCI = NA,
                          upperCI = NA
)

stats.rep.IV.interaction.lm = data.frame(Fvalue = sprintf("%.2f", result.lm$`F value`[4]),
                          df1 = result.lm$Df[4],
                          df2 = result.lm$Df[5],
                          pvalue = formatPval(result.lm$`Pr(>F)`[4]),
                          MD = NA,
                          lowerCI = NA,
                          upperCI = NA
)

stats.rep.CV.lm = data.frame(Fvalue = sprintf("%.2f", result.lm$`F value`[1]),
                          df1 = result.lm$Df[1],
                          df2 = result.lm$Df[5],
                          pvalue = formatPval(result.lm$`Pr(>F)`[1]),
                          MD = NA,
                          lowerCI = NA,
                          upperCI = NA
)
```

# Comparing ANCOVA in original study with reanalysis
## Independent variables

Main effect irrigation
```{r}
tab.IV = rbind(stats.orig.IV.irrigation, stats.rep.IV.irrigation, stats.rep.IV.irrigation.lm)
rownames(tab.IV) = c("original Study", "reanalysis type 3 SS", "reanalysis type 1 SS")
print(t(tab.IV))
```

Main effect composition
```{r}
tab.IV = rbind(stats.orig.IV.composition, stats.rep.IV.composition, stats.rep.IV.composition.lm)
rownames(tab.IV) = c("original Study", "reanalysis type 3 SS", "reanalysis type 1 SS")
print(t(tab.IV))
```

Interaction
```{r}
tab.IV = rbind(stats.orig.IV.interaction, stats.rep.IV.interaction, stats.rep.IV.interaction.lm)
rownames(tab.IV) = c("original Study", "reanalysis type 3 SS", "reanalysis type 1 SS")
print(t(tab.IV))
```

## Covariate
```{r}
tab.CV = rbind(stats.orig.CV, stats.rep.CV, stats.rep.CV.lm)
rownames(tab.CV) = c("original Study", "reanalysis type 3 SS", "reanalysis type 1 SS")
print(t(tab.CV))
```

# Assumptions
## 1.  Homogeneity of variance
* ANOVA/ANCOVA is fairly robust in terms of the error rate when sample sizes are equal.
* When groups with larger sample sizes have larger variances than the groups with smaller sample sizes, the resulting F-ratio tends to be conservative. That is, it's more likely to produce a non-significant result when a genuine difference does exist in the population.
* Conversely, when the groups with larger sample sizes have smaller variances than the groups with smaller samples sizes, the resulting F-ratio tends to be liberal and can inflate the false positive rate.
```{r}
tapply(dcl$Consumption, dcl$Irrigation, sd)
leveneTest(Consumption ~ Irrigation, data = dcl)
```
```{r}
tapply(dcl$Consumption, dcl$Composition, sd)
leveneTest(Consumption ~ Composition, data = dcl)
```

## 2.  Independence between covariate and IV
When the covariate and the experimental effect (independent variable) are not independent the treatment effect is obscured, spurious treatment effects can arise and the interpretation of the ANCOVA is seriously compromised.

We test whether our groups differ on the CV. If the groups do not significantly differ then is appropriate to use the covariate.
```{r}
fit.cv = aov(WCi ~ Irrigation, data = dcl) 
Anova(fit.cv, type=3)
```
```{r}
fit.cv = aov(WCi ~ Composition, data = dcl) 
Anova(fit.cv, type=3)
```

## 3. Homogeneity of regression slopes
```{r}
fit.hrs = aov(Consumption ~ WCi*Irrigation, data = dcl)
Anova(fit.hrs, type=3)
```

```{r fig.height=2, fig.width=3.5}
ggplot(dcl, aes(y=Consumption, x=WCi, color=Irrigation, shape=Irrigation)) +
  geom_point() +
  geom_smooth(formula = y ~ x,method=lm, se=FALSE, fullrange=TRUE) +
  theme_minimal() +
  theme(legend.title = element_blank())
```
```{r}
fit.hrs = aov(Consumption ~ WCi*Composition, data = dcl)
Anova(fit.hrs, type=3)
```

```{r fig.height=2, fig.width=3.5}
ggplot(dcl, aes(y=Consumption, x=WCi, color=Composition, shape=Composition)) +
  geom_point() +
  geom_smooth(formula = y ~ x,method=lm, se=FALSE, fullrange=TRUE) +
  theme_minimal() +
  theme(legend.title = element_blank())
```

# Notes

* We could reproduce the first reported ANCOVA, but had to exclude a single outlier in the covariate which was not clearly declared in the methods section.
* Model was clearly specified and F-values, p-values and Dfs were all reported.
* Type I SS was used, but no large differences compared to type 3 SS.
* Altogether 5 outcome variables and 5 ANCOVAs performed, but not adjusted for multiplicity.
* Assumptions of homogeneity of variances was met.
* Assumption of independence of IV and CV was only met for one IV but not for the other.
* Assumption of homogeneity of regression slopes was met (also thanks to the exclusion of the outlier!).

Data was analyzed according to recommendations by Field, Miles, & Field (2012).