---
title: "Reanalysis of 24-Fujiwara"
author: 
- Audrey Yeo^[University of Zurich, audrey.yeo@uzh.ch]
- Simon Schwab^[University of Zurich, simon.schwab@uzh.ch]
date: "01/06/2020"
output: pdf_document
toc: true
---

# Reference
Fujiwara, E. (2018). Looking at the eyes interferes with facial emotion recognition in alexithymia. Journal of Abnormal Psychology, 127(6), 571–577. https://doi.org/10.1037/abn0000361

```{r, include=FALSE}
# install.packages("biostatUZH", repos="http://R-Forge.R-project.org")
library(biostatUZH)
library(car)
library(readxl)
library(foreign)
library(readxl)
library(ggplot2)
library(cowplot)
library(ez)
```

# Notes from reading methods section
This study assessed the two Alexithymic groups, high and low, with the individuals' ability to identify clear and ambiguous blends of emotions in faces and their underlying visual attention patterns. Using the Toronto Alexithymia Scale, students with high alexithymia (HA, n = 73) or low alexithymia (LA, n = 76) were enrolled in this study. 

* Dependant variable: deviance score
* Independent variables: 
  * 2 levels ambiguity (clear vs. ambiguous emotional blends in the target face)
  * 6 levels of emotion (anger, disgust, fear, happiness, sadness, surprise is dominant emotion)
  * 2 groups (HA, LA; high and low Alexithymia score)
* Covariate:
  * DASS-21 (Depression, Anxiety and Stress total score)
  * Maybe log transformed response times were also used as covariate
* Design: 2 (w) x 6 (w) x 2 (b) mixed ANCOVA (w within; b between)

# Reading data
Data is loaded, reshaped into long form, and factors are specified.
```{r, echo = FALSE}
PATH = file.path(path.expand("~"), "Data", "ancova") # ancova project folder
data.orig = read_excel(file.path(PATH, "dataPrimaryStudies", "24-Fujiwara", "FujiwaraJAbnorm2019_data.xlsx"))

# convert to long format
N = 73 + 76
k = 6 * 2
data = data.frame(id = rep(1:N, 12))
data$dev = c(data.matrix(data.orig[,4:15]))
data$emo = c(rep("anger", N), rep("disgust", N), rep("fear", N), 
             rep("happiness", N), rep( "sadness", N), rep("surprise", N))
data$amb = c(rep("clear", N*6), rep("ambigious", N*6))
data$TAS_group = rep(data.orig$TAS_group, k)
data$DASS_tot = rep(data.orig$DASS_tot, k)
data$log_lat = rep(data.orig$log_lat, k)

# add group variable which is more informative
data$group = NA
data$group[data$TAS_group == 1] = "Low Alexithymia"
data$group[data$TAS_group == 3] = "High Alexithymia"

data$id = as.factor(data$id)
data$emo = as.factor(data$emo)
data$amb = as.factor(data$amb)
data$group = as.factor(data$group)

# order level of emotions from low to high avergae ambiguity
tmp = aggregate(dev ~ emo, data = data, FUN = mean)
myLevels = tmp$emo[order(tmp$dev)]
data$emo = factor(data$emo, levels = myLevels)

summary(data)
```

# Descriptives
## Dependant variable
Number of samples and mean (SD) in levels of the independent variables. We reproduce Table 3 and Figure 2A of the study.
```{r fig.height=2, fig.width=11}
d.1 = aggregate(dev ~ DASS_tot + log_lat + id*group, data = data, FUN = mean)
d.2 = aggregate(dev ~ DASS_tot + log_lat + id*emo, data = data, FUN = mean)
d.3 = aggregate(dev ~ DASS_tot + log_lat + id*amb, data = data, FUN = mean)

p1 = ggplot(d.1, aes(y=dev, x=group, color=group)) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.15, height = 0)) +
  theme_minimal() + theme(axis.text.x = element_blank())

p2 = ggplot(d.2, aes(y=dev, x=emo, color=emo)) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.15, height = 0)) +
  theme_minimal() + theme(axis.text.x = element_blank())

p3 = ggplot(d.3, aes(y=dev, x=amb, color=amb)) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.15, height = 0)) +
 theme_minimal() + theme(axis.text.x = element_blank())

plot_grid(p1, p2, p3, nrow = 1, ncol = 3)
```
Averaging subjects across along the third factor to see the two-way interactions. Plots show means (95%-CI as error bar)
```{r fig.height=2, fig.width=11}
d.A = aggregate(dev ~ group*emo, data = data, FUN = mean)
d.B = aggregate(dev ~ group*amb, data = data, FUN = mean)
d.C = aggregate(dev ~ emo*amb, data = data, FUN = mean)

# get SD
d.A$sd = aggregate(dev ~ group*emo, data = data, FUN = sd)[,3]
d.A$se = d.A$sd/sqrt(rep(c(73, 76), 6))

d.B$sd = aggregate(dev ~ group*amb, data = data, FUN = sd)[,3]
d.B$se = d.B$sd/sqrt(rep(c(73, 76), 2))

d.C$sd = aggregate(dev ~ emo*amb, data = data, FUN = sd)[,3]
d.C$se = d.C$sd/sqrt(rep(N, 12))

pA = ggplot(d.A, aes(y=dev, x=emo, group=group, color=group)) +
  geom_errorbar(aes(ymin=dev-1.96*se, ymax=dev+1.96*se), width=.2) +
  geom_line() + geom_point() +
  theme_minimal() + theme(axis.text.x = element_text(angle = 20) )

pB = ggplot(d.B, aes(y=dev, x=amb, group=group, color=group)) +
  geom_errorbar(aes(ymin=dev-1.96*se, ymax=dev+1.96*se), width=.2) +
  geom_line() + geom_point() +
  theme_minimal()

pC = ggplot(d.C, aes(y=dev, x=emo, group=amb, color=amb)) +
  geom_errorbar(aes(ymin=dev-1.96*se, ymax=dev+1.96*se), width=.2) +
  geom_line() + geom_point() +
  theme_minimal() + theme(axis.text.x = element_text(angle = 20))

plot_grid(pA, pB, pC, nrow = 1, ncol = 3)
```

## Covariate(s)
The two covariates are DASS_tot scores and log of time.
```{r fig.height=2, fig.width=8}
p1 = ggplot(d.1, aes(y=DASS_tot, x=group, color=group)) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.15, height = 0)) +
 theme_minimal() + theme(axis.text.x = element_blank())

p2 = ggplot(d.1, aes(y=log_lat, x=group, color=group)) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.15, height = 0)) +
 theme_minimal() + theme(axis.text.x = element_blank())

plot_grid(p1, p2, nrow = 1, ncol = 2)
```

# Main analysis ANCOVA
Independent variable: TAS_group (Between Group)
```{r, include=FALSE}
stats.orig.IV.amb = data.frame(
  Fvalue =  4.67,
  df1 = 1,
  df2 = 145,
  pvalue = 0.03)

stats.orig.IV.emoXamb = data.frame(
  Fvalue =  2.61,
  df1 = 4.21,
  df2 = 610.65,
  pvalue = 0.03)
```

```{r}
# Orthogonal contrasts
contrasts(data$group) = contr.helmert(2)
contrasts(data$amb) = contr.helmert(2) 
contrasts(data$emo) = contr.helmert(6) 

fit.ancova = aov(dev ~ log_lat + DASS_tot + (group*emo*amb) + Error(id/(emo*amb)) + group, 
                 data = data)
result = summary(fit.ancova) # Type I
print(result)
```

```{r}
fit2.ancova = ezANOVA(dv = .(dev), wid = .(id), between = .(group), within = .(emo, amb),
                      within_covariates = .(DASS_tot, log_lat), type=3, detailed=TRUE,  data = data)
print(fit2.ancova)
```

```{r, include=FALSE}
stats.rep.IV.amb = data.frame(Fvalue = sprintf("%.2f",result$`Error: id:amb`[[1]]$`F value`[1]),
                              df1 = result$`Error: id:amb`[[1]]$Df[1],
                              df2 = result$`Error: id:amb`[[1]]$Df[3],
                              pvalue = formatPval(result$`Error: id:amb`[[1]]$`Pr(>F)`[1])
)

stats.rep.IV.emoXamb = data.frame(Fvalue = sprintf("%.2f",result$`Error: id:emo:amb`[[1]]$`F value`[1]),
                                  df1 = result$`Error: id:emo:amb`[[1]]$Df[1],
                                  df2 = result$`Error: id:emo:amb`[[1]]$Df[3],
                                  pvalue = formatPval(result$`Error: id:emo:amb`[[1]]$`Pr(>F)`[1])
)

# ezANOVA with type 3 SS
stats.rep2.IV.amb = data.frame(Fvalue = sprintf("%.2f",fit2.ancova$ANOVA$F[4]),
                              df1 = fit2.ancova$ANOVA$DFn[4],
                              df2 = fit2.ancova$ANOVA$DFd[4],
                              pvalue = formatPval(fit2.ancova$ANOVA$p[4])
)

stats.rep2.IV.emoXamb = data.frame(Fvalue = sprintf("%.2f",fit2.ancova$ANOVA$F[7]),
                              df1 = fit2.ancova$ANOVA$DFn[7],
                              df2 = fit2.ancova$ANOVA$DFd[7],
                              pvalue = formatPval(fit2.ancova$ANOVA$p[7])
)
```

# Comparing ANCOVA in original study with reanalysis
## Independant variable

Main effect ambiguity
```{r}
tab.IV = rbind(stats.orig.IV.amb, stats.rep.IV.amb, stats.rep2.IV.amb)
rownames(tab.IV) = c("original Study", "reanalysis type I SS", "reanalysis type III SS")
print(t(tab.IV))
```

Interaction ambiguity x emotion
```{r}
tab.IV = rbind(stats.orig.IV.emoXamb, stats.rep.IV.emoXamb, stats.rep2.IV.emoXamb)
rownames(tab.IV) = c("original Study", "reanalysis type I SS", "reanalysis type III SS")
print(t(tab.IV))
```

# Assumptions
## 1.  Homogeneity of variance
* ANOVA/ANCOVA is fairly robust in terms of the error rate when sample sizes are equal.
* When groups with larger sample sizes have larger variances than the groups with smaller sample sizes, the resulting F-ratio tends to be conservative. That is, it's more likely to produce a non-significant result when a genuine difference does exist in the population.
* Conversely, when the groups with larger sample sizes have smaller variances than the groups with smaller samples sizes, the resulting F-ratio tends to be liberal and can inflate the false positive rate.
```{r}
tapply(d.1$dev, d.1$group, sd)
leveneTest(dev ~ group, data = d.1)
```


## 2.  Independence between covariate and IV.
When the covariate and the experimental effect (independent variable) are not independent the treatment effect is obscured, spurious treatment effects can arise and the interpretation of the ANCOVA is seriously compromised.

We test whether our groups differ on the CV. If the groups do not significantly differ then is appropriate to use the covariate.
```{r}
fit.cv1 = aov(DASS_tot ~ group, data = d.1) 
summary(fit.cv1)

fit.cv2 = aov(log_lat ~ group, data = d.1) 
summary(fit.cv2)
```


## 3. Homogeneity of regression slopes
```{r}
fit.hrs = aov(dev ~ DASS_tot*group, data = d.1)
Anova(fit.hrs, type = "III") 
```

```{r fig.height=2, fig.width=3.5}
ggplot(d.1, aes(y=dev, x=DASS_tot, color=group, shape=group)) +
  geom_point() +
  geom_smooth(formula = y ~ x,method=lm, se=FALSE, fullrange=TRUE) +
  theme_minimal() +
  theme(legend.title = element_blank())
```

```{r}
fit.hrs = aov(dev ~ DASS_tot*emo + Error(id/emo), data = d.2)
summary(fit.hrs)
```

```{r fig.height=2, fig.width=3.5}
ggplot(d.2, aes(y=dev, x=DASS_tot, color=emo, shape=emo)) +
  geom_point() +
  geom_smooth(formula = y ~ x,method=lm, se=FALSE, fullrange=TRUE) +
  theme_minimal() +
  theme(legend.title = element_blank())
```

```{r}
fit.hrs = aov(dev ~ DASS_tot*amb + Error(id/amb), data = d.3)
summary(fit.hrs)
```

```{r fig.height=2, fig.width=3.5}
ggplot(d.3, aes(y=dev, x=DASS_tot, color=amb, shape=amb)) +
  geom_point() +
  geom_smooth(formula = y ~ x,method=lm, se=FALSE, fullrange=TRUE) +
  theme_minimal() +
  theme(legend.title = element_blank())
```

# Notes
* Log transform of outcome could have been considered.
* Not all main effects and interactions were reported.
* Could not numerically reproduce the two reported results, but conclusions remain the same.
* Homogeneity of variances was met for groups.
* Independence between covariate DASS_tot and the IV group was not met.
* Homogeneity of regression slopes was mostly met, i.e. for group and emotion, but not across levels of ambiguity (but only weak evidence of violation).

Data was analyzed according to recommendations by Field, Miles, & Field (2012).