---
title: "24-Fujiwara"
author: "Audrey Yeo^[University of Zurich, audrey.yeo@uzh.ch]"
date: "01/06/2020"
output: pdf_document
toc: true
---

# Reference
Fujiwara, E. (2018). Looking at the eyes interferes with facial emotion recognition in alexithymia. *Journal of Abnormal Psychology*, 127(6), 571???577. https://doi.org/10.1037/abn0000361

```{r, include=FALSE}
# install.packages("biostatUZH", repos="http://R-Forge.R-project.org")
library(biostatUZH)
library(foreign) # for spss
library(ggplot2)
library(car) # for Anova
library(cowplot)
```

# Notes from reading methods section
This study assessed the two Alexithymic groups, high and low, with the individuals' ability to identify clear and ambiguous blends of emotions in faces and their underlying visual attention patterns. Using the Toronto Alexithymia Scale, students with high alexithymia (HA, n = 73) or low alexithymia (LA, n = 76) were enrolled in this study. 
* Dependant variable: DEV (combination)
* Independant variable: TAS_group (high or low Alexithymia score), Emotion (6 levels), Ambiguity (2 levels)
* Covariate: 
  * DASS_tot (Depression, Anxiety and Stress total scores or DASS-21)
  * log_lat (log transformed response time)

# Reading data
Data is loaded, reshaped into long form, and factors are specified.
```{r, echo = FALSE}
# PATH = file.path(path.expand("~"), "Desktop", "CRS") # ancova project folder
# PATH1 = file.path(PATH, "dataPrimaryStudies", "24-Fujiwara", "24-Fujiwara-1.sav")
# spss_data <- file.choose(PATH1)
# spss_data <- read.spss(spss_data, to.data.frame = TRUE) # external window opens for this
# data <- spss_data
# save(data, file = "../dataPrimaryStudies/24-Fujiwara/24-Fujiwara.rda") # easier to call
load("../../ancova/dataPrimaryStudies/24-Fujiwara/24-Fujiwara.rda")
#long = data.matrix(data[, 4:15])
#var_names = t(colnames(data))[4:15] # there are the column names I should have 2x of
longdata = data.frame(dev = c(data.matrix(data[,4:15])))
#longdata$dev = data.frame(longdata$dev)
longdata$id = as.factor(rep(1:149, 12))
longdata$TAS_group = as.factor(rep(data$TAS_group, 12))
longdata$TAS_group = factor(longdata$TAS_group, levels = c("No Alexithymia", "High Alexithymia"), ordered =FALSE)
longdata$amb = as.factor(c(rep("clear", 149*6), rep("ambigious", 149*6)))
longdata$emo = as.factor(c(rep("anger", 149), rep("disgust",149), rep("fear",149),
                           rep("happiness", 149), rep( "sadness", 149), rep("surprise", 149)))
longdata$dass = rep(data$DASS_tot, 12)
longdata$time = rep(data$log_lat, 12)
```

# Descriptives
## Dependant variable
Number of samples and mean (SD) in levels of the independant variables. We reproduce Table 3 and Figure 2A of the study.
```{r}
idx = c(1, 3, 2, 4) # sorting as in publication
tab.dv = array(NA, dim=c(2,3))
tab.dv[,1] = levels(longdata$TAS_group)
tab.dv[,2] = summary(longdata$TAS_group)
tab.dv[,3] = tapply(longdata$dev, longdata$TAS_group, 
                  function (x) sprintf("%0.2f (%0.2f)", mean(x), sd(x)))
colnames(tab.dv) = c("group", "n", "mean (SD)")
print(tab.dv)
```

```{r fig.height=2, fig.width=3.5}
ggplot(longdata, aes(y=dev, x=TAS_group, color=group)) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.15, height = 0)) +
  theme_minimal() +
  theme(axis.text.x = element_blank(), legend.title = element_blank()) +
  xlab("group") + ylab("anger score") + ggtitle("Dependant variable")
```

## Covariate(s)
The two covariates are DASS_tot scores and log of time.
```{r}
#idx = c(1, 3, 2, 4) # sorting as in publication
tab.cv = array(NA, dim=c(2,3))
tab.cv[,1] = levels(longdata$TAS_group)
tab.cv[,2] = summary(longdata$TAS_group)
tab.cv[,3] = tapply(longdata$dass, longdata$TAS_group, 
                  function (x) sprintf("%0.1f (%0.1f)", mean(x), sd(x)))
colnames(tab.cv) = c("group", "n", "mean of DASS (SD)")
tab.cv1 = tab.cv
print(tab.cv1)
```

```{r fig.height=2, fig.width=3.5, echo = FALSE}
p1 = ggplot(longdata, aes(y=dass, x=TAS_group, color=TAS_group)) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.15, height = 0)) +
  theme_minimal() +
  theme(axis.text.x = element_blank(), legend.title = element_blank()) +
  xlab("group") + ylab("DASS scores") + ggtitle("Covariate")
p2 = ggplot(longdata, aes(y=time, x=TAS_group, color=TAS_group)) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.15, height = 0)) +
  theme_minimal() +
  theme(axis.text.x = element_blank(), legend.title = element_blank()) +
  xlab("group") + ylab("log time (millisec)") + ggtitle("Covariate")
plot_grid(p1, p2, nrow = 1, ncol = 2)
```
Covariate log of time's mean and standard deviation
```{r}
idx = c(1, 3, 2, 4) # sorting as in publication
tab.cv = array(NA, dim=c(2,3))
tab.cv[,1] = levels(longdata$TAS_group)
tab.cv[,2] = summary(longdata$TAS_group)
tab.cv[,3] = tapply(longdata$time, longdata$TAS_group, 
                  function (x) sprintf("%0.1f (%0.1f)", mean(x), sd(x)))
colnames(tab.cv) = c("group", "n", "mean of log time (SD)")
tab.cv2 = tab.cv
print(tab.cv2)
```

# Main analysis ANCOVA
Independant variable: Ambiguity
```{r, include=FALSE}
stats.orig.IV1 = data.frame(
  Fvalue =  33.07,
  df1 = 1,
  df2 = 146,
  pvalue = "<0.001")

stats.orig.CV1 = stats.orig.IV1
stats.orig.CV2 = stats.orig.IV1
```

Independant variable: Emotion
```{r, include=FALSE}
stats.orig.IV2 = data.frame(
  Fvalue =  4.75,
  df1 = 4.72,
  df2 = 689.06,
  pvalue = "<0.001")
```



```{r, include=FALSE}
stats.orig.CV1 = data.frame(
  Fvalue =  33.07,
  df1 = 1,
  df2 = 146,
  pvalue = "<0.001")

stats.orig.CV2 = data.frame(
  Fvalue =  33.07,
  df1 = 1,
  df2 = 146,
  pvalue = "<0.001")
```


```{r}
# Orthogonal contrasts
contrasts(longdata$TAS_group) = contr.helmert(2) 
contrasts(longdata$amb) = contr.helmert(2) 
contrasts(longdata$emo) = contr.helmert(6) 

aov(dev ~ time + dass + group*emo*amb + Error(id/(emo*amb)) + group, data = longdata)


fit.ancova = aov(dev ~ time + dass + group*emo*amb, data = longdata)
result = Anova(fit.ancova)


# fit.ancova = aov(dev ~ time + dass + group*emo*amb + group, data = longdata)
# Anova(fit.ancova, type = "III")
```

```{r, include=FALSE}
stats.rep.IV1 = data.frame(Fvalue = sprintf("%.2f",result$`F value`[3]),
                          df1 = result$Df[3],
                          df2 = result$Df[6],
                          pvalue = formatPval(result$`Pr(>F)`[3]))

stats.rep.IV2 = data.frame(Fvalue = sprintf("%.2f",result$`F value`[3]),
                          df1 = result$Df[4],
                          df2 = result$Df[6],
                          pvalue = formatPval(result$`Pr(>F)`[3]))

stats.rep.CV1 = data.frame(Fvalue = sprintf("%.2f",result$`F value`[2]),
                          df1 = result$Df[1],
                          df2 = result$Df[6],
                          pvalue = formatPval(result$`Pr(>F)`[2]))

stats.rep.CV2 = data.frame(Fvalue = sprintf("%.2f",result$`F value`[2]),
                          df1 = result$Df[2],
                          df2 = result$Df[6],
                          pvalue = formatPval(result$`Pr(>F)`[2]))
```

# Comparing ANCOVA in original study with reanalysis
## Independant variable
```{r}
tab.IV = rbind(stats.orig.IV1, stats.orig.IV1,  stats.rep.IV2)
rownames(tab.IV) = c("original study's IV: Ambiguity", "original study's IV: Emotion", "reanalysis")
print(t(tab.IV))
```

## Covariate
```{r}
tab.CV = rbind(stats.orig.CV1, stats.orig.CV2, stats.rep.CV1, stats.rep.CV2)
rownames(tab.CV) = c("original Study COV : Time", "original Study COV: DASS", "reanalysis of COV : Time", "reanalysis of COV : DASS")
print(t(tab.CV))
```

# Assumptions
## 1.  Homogeneity of variance
* ANOVA/ANCOVA is fairly robust in terms of the error rate when sample sizes are equal.
* When groups with larger sample sizes have larger variances than the groups with smaller sample sizes, the resulting F-ratio tends to be conservative. That is, it's more likely to produce a non-significant result when a genuine difference does exist in the population.
* Conversely, when the groups with larger sample sizes have smaller variances than the groups with smaller samples sizes, the resulting F-ratio tends to be liberal and can inflate the false positive rate.

There is a similar standard deviation between groups. There is no evidence of significance that there is no homogeneity of variance.
```{r}
tapply(longdata$dass, longdata$TAS_group, sd)
leveneTest(dass ~ TAS_group, data = longdata)
```


## 2.  Independence between covariate and IV
When the covariate and the experimental effect (independent variable) are not independent the treatment effect is obscured, spurious treatment effects can arise and the interpretation of the ANCOVA is seriously compromised.

We test whether our groups differ on the CV. If the groups do not significantly differ then is appropriate to use the covariate.

There is no evidence of significance, that groups differ in their DASS_total scores but marginal evidence that groups differ with log of time). Thus, these covariate is appropriate for this analysis.
```{r}
fit.cv1 = aov(longdata$dass ~ TAS_group, data = longdata) 
summary(fit.cv1)

fit.cv2 = aov(longdata$time ~ TAS_group, data = longdata) 
summary(fit.cv2)
```


## 3. Homogeneity of regression slopes

There is no evidence of significance that groups interact with DASS scores, neither does group interact with time, this is confirmed on visual inspection.
```{r}
fit.hrs = aov(dev ~ dass*TAS_group, data = longdata)
Anova(fit.hrs, type = "II")

fit.hrs = aov(dev ~ time*TAS_group, data = longdata)
Anova(fit.hrs, type="II")
```

```{r fig.height=2, fig.width=3.5}
p1 = ggplot(longdata, aes(y=dev, x=time, color = TAS_group, shape= TAS_group)) +
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm, se=FALSE, fullrange=TRUE) +
  theme_minimal() +
  theme(legend.title = element_blank()) +
  xlab("Time") + ylab("Deviance scores")

p2 = ggplot(longdata, aes(y=dev, x=dass, color = TAS_group, shape= TAS_group)) +
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm, se=FALSE, fullrange=TRUE) +
  theme_minimal() +
  theme(legend.title = element_blank()) +
  xlab("DASS-21 score") + ylab("Deviance scores")

plot_grid(p1, p2, nrow = 1, ncol = 2)
```


# Notes

<!-- * The first reported ANCOVA which was n.s. was reproduced --> -->
* Assumptions were met for ancova analysis.

Data was analyzed according to recommendations by Field, Miles, & Field (2012).