---
title: "25-VanLeijsen"
author: "Simon Schwab, Audrey Yeo"
date: "11/13/2019"
output: pdf_document
---

# Reference
Van Leijsen et al. (2017). Plasma AB (Amyloid-B) Levels and Severity and Progression of Small Vessel Disease. Stroke, 16(5), 351–359. https://doi.org/10.1161/STROKEAHA.117.019810

We first load the appropriate packages
```{r, eval=TRUE, echo=FALSE, include = FALSE}
#Setting up libraries
knitr::opts_chunk$set(echo = TRUE)
library(biostatUZH)
library(car)
library(readxl)
library(foreign)
library(readxl)
library(ggplot2)
library(tidyr)
library(beeswarm)
```

# Notes from reading methods section

* Dependant variable: AB38
* The independant variables 
  * 1 or more microbleeds (n = 81) 
  * no microbleed (n=405)
  * 1 or more lacunes (n = 132)
  * no lacunes (n = 355)
* Covariate: age, sex and hypertension

```{r}
stats.orig.IV = data.frame(
  Fvalue =  NA,
  df1 = NA,
  df2 = NA,
  pvalue = 0.01, # for microbleed presence and for lacune presence groups
  MD = NA,
  lowerCI = NA,
  upperCI = NA)

stats.orig.CV = stats.orig.IV
```

# Reading data
Data is loaded, reshaped if necessary, and factors are specified.
```{r}
data = read_excel("../results/data/25-VanLeijsen/VanLeijsen-1.xlsx")
data = data.frame(data)
str(data)
data$mb_presence_b = data$mb_presence_b == 1
data$lac_presence_b = data$lac_presence_b == 1
data$sex.factor = NA
data$sex.factor[data$sex == 1] = "female"
data$sex[data$sex == 2] = "0"
data$sex.factor[data$sex == 0] = "male"
data$hypertension.factor = NA
data$hypertension  = data$hypertension == 1 

data$Group.factor = NA # only used for table
data[data$mb_presence_b == TRUE,]$Group.factor = "microbleed"
data[data$mb_presence_b == FALSE,]$Group.factor = "no microbleed"
data[data$lac_presence_b == TRUE,]$Group.factor = "lacunes"
data[data$lac_presence_b == FALSE,]$Group.factor = "no lacunes"
unique(data$Group.factor)
data$Group.factor = factor(data$Group.factor,
                           levels = c("no microbleed", "microbleed",
                                      "no lacunes", "lacunes"))
```

# Descriptives 
Number of samples and mean (SD) in levels of the independant variables. We reproduce the mean and sd values of Table 2 of this study
```{r}
a = sprintf("%.1f (%.1f)", mean(data$AB38[data$mb_presence_b == 0]), sd(data$AB38[data$mb_presence_b == 0]))
b = sprintf("%.1f (%.1f)", mean(data$AB38[data$mb_presence_b == 1]), sd(data$AB38[data$mb_presence_b == 1]))

c = sprintf("%.1f (%.1f)", mean(data$AB38[data$lac_presence_b == 0]), sd(data$AB38[data$lac_presence_b == 0]))
d = sprintf("%.1f (%.1f)", mean(data$AB38[data$lac_presence_b == 1]), sd(data$AB38[data$lac_presence_b == 1]))

idx = c(1, 3, 2, 4) # sorting as in publication
tab.dv = array(NA, dim=c(4,3))
tab.dv[,1] = levels(data$Group.factor)
tab.dv[,2] = c("405", "81", "355", "132")
tab.dv[,3] = c(a, b, c, d)
colnames(tab.dv) = c("group", "n", "mean (SD)")
print(tab.dv)
```
# boxplot with DV
* Upon visual inspection, each level of each independant group seem to have a similar outcome effect.
```{r, fig.height=3, fig.width=4.5}
#IV : mb_presence_b
ggplot(data, aes(y=AB38, x = mb_presence_b, color =mb_presence_b )) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.15, height = 0)) +
  theme_minimal() +
  theme(axis.text.x = element_blank(), legend.title = element_blank()) +
  xlab("Microbleed Presence") + ylab("AB38") + 
  ggtitle("Dependant variable with Microbleed presence")
#IV : lac_presence_b
ggplot(data, aes(y=AB38, x = lac_presence_b, color =lac_presence_b )) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.15, height = 0)) +
  theme_minimal() +
  theme(axis.text.x = element_blank(), legend.title = element_blank()) +
  xlab("Lacune Presence") + ylab("AB38") + 
  ggtitle("Dependant variable with Lacune presence")
```
  
# Descriptives 
## COV with boxplot
```{r}
e = sprintf("%.1f (%.1f)", mean(data$age[data$mb_presence_b == 0]), 
            sd(data$age[data$mb_presence_b == 0]))
f = sprintf("%.1f (%.1f)", mean(data$age[data$mb_presence_b == 1]), 
            sd(data$age[data$mb_presence_b == 1]))
g = sprintf("%.1f (%.1f)", mean(data$age[data$lac_presence_b == 0]), sd(data$age[data$lac_presence_b == 0]))
h = sprintf("%.1f (%.1f)", mean(data$age[data$lac_presence_b == 1]), sd(data$age[data$lac_presence_b == 1]))

idx = c(1, 3, 2, 4) # sorting as in publication
tab.cv = array(NA, dim=c(4,3))
tab.cv[,1] = levels(data$Group.factor)
tab.cv[,2] = c("405", "81", "355" , "132")
tab.cv[,3] = c(e,f,g,h)
colnames(tab.cv) = c("group", "n", "mean (SD)")
print(tab.cv)
```
* Age upon visual inspection is similar between groups
```{r, fig.height=3, fig.width=4.5}
ggplot(data, 
       aes(y=age, x=mb_presence_b, color=mb_presence_b)) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.15, height = 0)) +
  theme(legend.title = element_blank()) + theme_minimal() + 
  theme(axis.text.x = element_blank(), legend.title = element_blank()) +
  labs(x = "Lacune presence", y = "Age", title = "" )
ggplot(data, 
       aes(y=age, x=lac_presence_b, color=lac_presence_b)) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.15, height = 0)) +
  theme(legend.title = element_blank()) + theme_minimal() + 
  theme(axis.text.x = element_blank(), legend.title = element_blank()) +
  labs(x = "Lacune presence", y = "Age", title = "" )
```

# Main analysis ANCOVA
```{r}
stats.orig.IV = data.frame(
  Fvalue =  NA,
  df1 = NA,
  df2 = NA,
  pvalue = NA,
  MD = NA,
  lowerCI = NA,
  upperCI = NA)

stats.orig.CV = stats.orig.IV

# enter all results from primary study here
stats.orig.IV$pvalue = 0.01
``` 

We verify the p values stated in Table 2 with respect to variables.
* There is evidence of significance that Age as a covariate contribute to variance in outcome (p = 8.037e-05)
```{r}
# Orthogonal contrasts
contrasts(data$mb_presence_b) = contr.helmert(2) 
contrasts(data$lac_presence_b) = contr.helmert(2) 

#contrasts(data$IV)<-cbind(c(-2,1,1), c(0,-1,1))
fit <- aov(AB38 ~ age + mb_presence_b + lac_presence_b + sex.factor + hypertension, data = data)
summary(fit)
result = Anova(fit, type = 3) # this is not a balanced study
print(result)
```

```{r}
stats.rep.IV = data.frame(Fvalue = sprintf("%.2f",result$`F value`[3]),
                          df1 = result$Df[3],
                          df2 = result$Df[4],
                          pvalue = formatPval(result$`Pr(>F)`[3]),
                          MD = NA,
                          lowerCI = NA,
                          upperCI = NA
)

stats.rep.CV = data.frame(Fvalue = sprintf("%.2f",result$`F value`[2]),
                          df1 = result$Df[2],
                          df2 = result$Df[4],
                          pvalue = formatPval(result$`Pr(>F)`[2]),
                          MD = NA,
                          lowerCI = NA,
                          upperCI = NA
)
```

# Comparing ANCOVA in original study with reanalysis
## Independant variable
```{r}
tab.IV = rbind(stats.orig.IV, stats.rep.IV)
rownames(tab.IV) = c("original Study", "reanalysis")
print(t(tab.IV))
```
## Covariate
```{r}
tab.CV = rbind(stats.orig.CV, stats.rep.CV)
rownames(tab.CV) = c("original Study", "reanalysis")
print(t(tab.CV))
```
# Assumptions
## 1.  Homogeneity of variance
* ANOVA/ANCOVA is fairly robust in terms of the error rate when sample sizes are equal.
* When groups with larger sample sizes have larger variances than the groups with smaller sample sizes, the resulting F-ratio tends to be conservative. That is, it's more likely to produce a non-significant result when a genuine difference does exist in the population.
* Conversely, when the groups with larger sample sizes have smaller variances than the groups with smaller samples sizes, the resulting F-ratio tends to be liberal and can inflate the false positive rate.
* In this study, there is homogeneity of variance
```{r}
tapply(data$AB38, data$mb_presence_b, sd)
tapply(data$AB38, data$lac_presence_b, sd)
leveneTest(AB38 ~ lac_presence_b, data = data) 
leveneTest(AB38 ~ mb_presence_b, data = data)
```
## 2.  Independence between covariate and IV
* The IV of this study have evidence of significant effect to the plasma level AB38.
```{r}
fit.cv = aov(AB38 ~ lac_presence_b + mb_presence_b, data = data) 
summary(fit.cv)
```

## 3. Homogeneity of regression slopes
* There is no evidence of significant interaction between covariate and independant variable. Thus we can assume homogeneity of regression slope
```{r}
fit.hrs = aov(AB38 ~ age*mb_presence_b, data = data)
Anova(fit.hrs, type=3) # no evidence of interaction, there is homogeneity of IV levels across age
fit.hrs = aov(AB38 ~ age*lac_presence_b, data = data)
Anova(fit.hrs, type=3) 
```

## Independance of Covariate with Independant Variables (Visual inspection of Homogeneity of Regression slopes)
* Visually, the two levels of each independant variable follows the same pattern - there seems to be independance of Covariate versus independant variable.
```{r fig.height=2, fig.width=3.5}
ggplot(data, aes(y=AB38, x= age, color= mb_presence_b)) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.15, height = 0)) +
  geom_smooth(formula = y ~ x,method=lm, se=FALSE, fullrange=TRUE) +
  theme_minimal() +
  theme(axis.text.x = element_blank(), legend.title = element_blank()) +
  xlab("Age") + ylab("AB38") + ggtitle("Covariate with Microbleed presence")

ggplot(data, aes(y=AB38, x= age, color= lac_presence_b)) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.15, height = 0)) +
  geom_smooth(formula = y ~ x,method=lm, se=FALSE, fullrange=TRUE) +
  theme_minimal() +
  theme(axis.text.x = element_blank(), legend.title = element_blank()) +
  xlab("Age") + ylab("AB38") + ggtitle("Covariate with Lacunes presence")
```

