---
title: "Reanalysis of 26-Keane"
author: 
- Simon Schwab^[University of Zurich, simon.schwab@uzh.ch]
date: "23/03/2021"
output: pdf_document
toc: true
---

# Reference
Keane, B. P., Paterno, D., Kastner, S., Krekelberg, B., & Silverstein, S. M. (2019). Intact illusory contour formation but equivalently impaired visual shape completion in first- and later-episode schizophrenia. Journal of Abnormal Psychology, 128(1), 57–68. https://doi.org/10.1037/abn0000384

```{r, include=FALSE}
# install.packages("biostatUZH", repos="http://R-Forge.R-project.org")
library(biostatUZH)
library(car)
library(readxl)
library(foreign)
library(readxl)
library(ggplot2)
library(cowplot)
library(ez)
library(nlme)
```

# Notes from reading methods section
This study tested three groups, healthy controls, first-espisode psychotic patients and chronic schizophrenia patients using a visual shape completition task. 
* Dependant variable: threshold values (log deg of rotation)
* Independent variables: 
  * relatability (illusory, fragmented)
  * contour type (traditional, wire)
  * patient group (control (n = 48), first episode (n = 23), chronic (n = 49) )
* Covariate:
  * Chlorpromazine equivalent does (CPZ)
* Design: 2 (w) x 2 (w) x 3 (b) mixed ANCOVA (w within; b between)

# Reading data
Data is loaded, reshaped into long form, and factors are specified.
```{r, echo = FALSE}
PATH = file.path(path.expand("~"), "Data", "ancova") # ancova project folder
data.orig = read.spss(file = file.path(PATH, "dataPrimaryStudies", "27-Keane",
                                       "VisualCompletionData_Final_Sep2018.sav"),
                      to.data.frame = TRUE)
data.orig = subset(data.orig, subset = grepl("^SF.*", subID)) # remove empty rows
data.orig = subset(data.orig, subset = !excludeThisSubj) # remove excluded subjects

# convert to long format
N  = 48 + 23 + 49
k = 2 * 2

data.orig$subID = trimws(data.orig$subID)
data.orig$Race  = trimws(data.orig$Race)

data = data.frame(id = as.factor(rep(data.orig$subID, k)))
data$age =  rep(data.orig$Age, k)
data$logDeg = c(data.matrix(data.orig[,6:9]))
data$relatability = as.factor(c(rep("illusory", N*2), rep("fragmented", N*2)))
# traditional vs wired, not clear what is NL und L
data$contour = as.factor(c(rep("NL", N), rep("L", N), rep("NL", N), rep("L", N)))
data$group = rep(data.orig$CSz, k)
data$CPZ = rep(data.orig$CPZ_New2, k)

summary(data)
```

# Descriptives
## Dependant variable
Number of samples and mean (SD) in levels of the independent variables. We reproduce Table 3 and Figure 2A of the study.
```{r fig.height=2, fig.width=11}
d.1 = aggregate(. ~ id*group, data = data, FUN = mean, na.rm=TRUE, na.action = "na.pass")
d.2 = aggregate(. ~ id*relatability, data = data, FUN = mean, na.rm=TRUE, na.action = "na.pass")
d.3 = aggregate(. ~ id*contour, data = data, FUN = mean, na.rm=TRUE, na.action = "na.pass")

p1 = ggplot(d.1, aes(y=logDeg, x=group, color=group)) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.15, height = 0)) +
  theme_minimal() + theme(axis.text.x = element_blank())

p2 = ggplot(d.2, aes(y=logDeg, x=relatability, color=relatability)) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.15, height = 0)) +
  theme_minimal() + theme(axis.text.x = element_blank())

p3 = ggplot(d.3, aes(y=logDeg, x=contour, color=contour)) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.15, height = 0)) +
 theme_minimal() + theme(axis.text.x = element_blank())

plot_grid(p1, p2, p3, nrow = 1, ncol = 3)
```
Averaging subjects across along the third factor to see the two-way interactions. Plots show means (95%-CI as error bar)
```{r fig.height=2, fig.width=11}
d.A = aggregate(logDeg ~ group*relatability, data = data, FUN = mean)
d.B = aggregate(logDeg ~ group*contour, data = data, FUN = mean)
d.C = aggregate(logDeg ~ relatability*contour, data = data, FUN = mean)

# get SD
d.A$sd = aggregate(logDeg ~ group*relatability, data = data, FUN = sd)[,3]
d.A$se = d.A$sd/sqrt(rep(c(48, 49, 23), 2))

d.B$sd = aggregate(logDeg ~ group*contour, data = data, FUN = sd)[,3]
d.B$se = d.B$sd/sqrt(rep(c(48, 49, 23), 2))

d.C$sd = aggregate(logDeg ~ relatability*contour, data = data, FUN = sd)[,3]
d.C$se = d.C$sd/sqrt(rep(N, 4))

pA = ggplot(d.A, aes(y=logDeg, x=relatability, group=group, color=group)) +
  geom_errorbar(aes(ymin=logDeg-1.96*se, ymax=logDeg+1.96*se), width=.2) +
  geom_line() + geom_point() +
  theme_minimal() + theme(axis.text.x = element_text(angle = 20) )

pB = ggplot(d.B, aes(y=logDeg, x=contour, group=group, color=group)) +
  geom_errorbar(aes(ymin=logDeg-1.96*se, ymax=logDeg+1.96*se), width=.2) +
  geom_line() + geom_point() +
  theme_minimal()

pC = ggplot(d.C, aes(y=logDeg, x=relatability, group=contour, color=contour)) +
  geom_errorbar(aes(ymin=logDeg-1.96*se, ymax=logDeg+1.96*se), width=.2) +
  geom_line() + geom_point() +
  theme_minimal() + theme(axis.text.x = element_text(angle = 20))

plot_grid(pA, pB, pC, nrow = 1, ncol = 3)
```

## Covariate(s)
The two covariates are DASS_tot scores and log of time.
```{r fig.height=2, fig.width=4}
ggplot(d.1, aes(y=CPZ, x=group, color=group)) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.15, height = 0)) +
  theme_minimal() + theme(axis.text.x = element_blank())
```

# Main analysis ANCOVA
Independent variable: TAS_group (Between Group)
```{r, include=FALSE}
stats.orig.IV.group = data.frame(
  Fvalue =  15.8,
  df1 = 2,
  df2 = 117,
  pvalue = "p < 0.001")

stats.orig.IV.cont = data.frame(
  Fvalue =  9.4,
  df1 = 1,
  df2 = 117,
  pvalue = 0.003)

stats.orig.IV.groupXrelat = data.frame(
  Fvalue = 6.1,
  df1 = 2,
  df2 = 117,
  pvalue = 0.003)

stats.orig.IV.relatXcont = data.frame(
  Fvalue = 13.2,
  df1 = 1,
  df2 = 117,
  pvalue = "p < 0.001")

stats.orig.IV.groupXrelatXcont = data.frame(
  Fvalue = 1.2,
  df1 = 2,
  df2 = 117,
  pvalue = 0.296)

stats.orig.CV = data.frame(
  Fvalue = NA,
  df1 = NA,
  df2 = NA,
  pvalue = "all ps > 0.60")
```

```{r}
# Orthogonal contrasts
contrasts(data$group) = contr.helmert(3)
contrasts(data$relatability) = contr.helmert(2) 
contrasts(data$contour) = contr.helmert(2) 
```

ANOVA was reported completely, but ANCOVA with a covariate only incompletely.
```{r}
fit.ancova = ezANOVA(dv = .(logDeg), wid = .(id), between = .(group), within = .(relatability, contour),
                     type=3, detailed=TRUE,  data = data)
print(fit.ancova)
```

# Checking interactions of the covariate CPZ with other variables using a linear mixed model as CPZ has inclompelte data.
```{r}
mod1 = lme(logDeg ~ CPZ + group*relatability*contour, data=data, na.action = na.exclude, random = ~1|id/relatability/contour, method = "ML")
result.lme = anova(mod1, type="marginal")
print(result.lme)
```

```{r, include=FALSE}
stats.rep.IV.group = data.frame(Fvalue = sprintf("%.2f",fit.ancova$ANOVA$F[2]),
                                df1 = fit.ancova$ANOVA$DFn[2],
                                df2 = fit.ancova$ANOVA$DFd[2],
                                pvalue = formatPval(fit.ancova$ANOVA$p[2])
)

stats.rep.IV.cont = data.frame(Fvalue = sprintf("%.2f",fit.ancova$ANOVA$F[4]),
                               df1 = fit.ancova$ANOVA$DFn[4],
                               df2 = fit.ancova$ANOVA$DFd[4],
                               pvalue = formatPval(fit.ancova$ANOVA$p[4])
)

stats.rep.IV.groupXrelat = data.frame(Fvalue = sprintf("%.2f",fit.ancova$ANOVA$F[5]),
                                      df1 = fit.ancova$ANOVA$DFn[5],
                                      df2 = fit.ancova$ANOVA$DFd[5],
                                      pvalue = formatPval(fit.ancova$ANOVA$p[5])
)

stats.rep.IV.relatXcont = data.frame(Fvalue = sprintf("%.2f",fit.ancova$ANOVA$F[7]),
                                     df1 = fit.ancova$ANOVA$DFn[7],
                                     df2 = fit.ancova$ANOVA$DFd[7],
                                     pvalue = formatPval(fit.ancova$ANOVA$p[7])
)

stats.rep.IV.groupXrelatXcont = data.frame(Fvalue = sprintf("%.2f",fit.ancova$ANOVA$F[8]),
                                           df1 = fit.ancova$ANOVA$DFn[8],
                                           df2 = fit.ancova$ANOVA$DFd[8],
                                           pvalue = formatPval(fit.ancova$ANOVA$p[8])
)

stats.rep.CV = data.frame(Fvalue = sprintf("%.2f",result.lme$`F-value`[2]),
                                df1 = result.lme$numDF[2],
                                df2 = result.lme$denDF[2],
                                pvalue = formatPval(result.lme$`p-value`[2])
)
```

# Comparing ANCOVA in original study with reanalysis
## Independant variable

Main effect group
```{r}
tab.IV = rbind(stats.orig.IV.group, stats.rep.IV.group)
rownames(tab.IV) = c("original Study", "reanalysis type III SS")
print(t(tab.IV))
```

Main effect contour
```{r}
tab.IV = rbind(stats.orig.IV.cont, stats.rep.IV.cont)
rownames(tab.IV) = c("original Study", "reanalysis type III SS")
print(t(tab.IV))
```

Interaction group X relatability
```{r}
tab.IV = rbind(stats.orig.IV.groupXrelat, stats.rep.IV.groupXrelat)
rownames(tab.IV) = c("original Study", "reanalysis type III SS")
print(t(tab.IV))
```

Interaction contour X relatability
```{r}
tab.IV = rbind(stats.orig.IV.relatXcont, stats.rep.IV.relatXcont)
rownames(tab.IV) = c("original Study", "reanalysis type III SS")
print(t(tab.IV))
```

Interaction group X relatability X contour
```{r}
tab.IV = rbind(stats.orig.IV.groupXrelatXcont, stats.rep.IV.groupXrelatXcont)
rownames(tab.IV) = c("original Study", "reanalysis type III SS")
print(t(tab.IV))
```

Covariate 
```{r}
tab.CV = rbind(stats.orig.CV, stats.rep.CV)
rownames(tab.CV) = c("original Study", "reanalysis type I SS")
print(t(tab.CV))
```

# Assumptions
## 1.  Homogeneity of variance
* ANOVA/ANCOVA is fairly robust in terms of the error rate when sample sizes are equal.
* When groups with larger sample sizes have larger variances than the groups with smaller sample sizes, the resulting F-ratio tends to be conservative. That is, it's more likely to produce a non-significant result when a genuine difference does exist in the population.
* Conversely, when the groups with larger sample sizes have smaller variances than the groups with smaller samples sizes, the resulting F-ratio tends to be liberal and can inflate the false positive rate.
```{r}
tapply(d.1$logDeg, d.1$group, sd)
leveneTest(logDeg ~ group, data = d.1)
```


## 2.  Independence between covariate and IV.
When the covariate and the experimental effect (independent variable) are not independent the treatment effect is obscured, spurious treatment effects can arise and the interpretation of the ANCOVA is seriously compromised.

We test whether our groups differ on the CV. If the groups do not significantly differ then is appropriate to use the covariate.
```{r}
fit.cv1 = aov(CPZ ~ group, data = d.1) 
summary(fit.cv1)
```


## 3. Homogeneity of regression slopes
```{r}
fit.hrs = aov(logDeg ~ CPZ*group, data = d.1)
Anova(fit.hrs, type = "III") 
```

```{r fig.height=2, fig.width=3.5}
ggplot(d.1, aes(y=logDeg, x=CPZ, color=group, shape=group)) +
  geom_point() +
  geom_smooth(formula = y ~ x,method=lm, se=FALSE, fullrange=TRUE) +
  theme_minimal() +
  theme(legend.title = element_blank())
```

```{r}
mod1 = lme(logDeg ~ CPZ + group*relatability, data=d.2, na.action = na.exclude, random = ~1|id/relatability, method = "ML")
result.lme = anova(mod1, type="marginal")
print(result.lme)
```

```{r fig.height=2, fig.width=3.5}
ggplot(d.2, aes(y=logDeg, x=CPZ, color=relatability, shape=relatability)) +
  geom_point() +
  geom_smooth(formula = y ~ x,method=lm, se=FALSE, fullrange=TRUE) +
  theme_minimal() +
  theme(legend.title = element_blank())
```

```{r}
mod1 = lme(logDeg ~ CPZ + group*contour, data=d.3, na.action = na.exclude, random = ~1|id/contour, method = "ML")
result.lme = anova(mod1, type="marginal")
print(result.lme)
```

```{r fig.height=2, fig.width=3.5}
ggplot(d.3, aes(y=logDeg, x=CPZ, color=contour, shape=contour)) +
  geom_point() +
  geom_smooth(formula = y ~ x,method=lm, se=FALSE, fullrange=TRUE) +
  theme_minimal() +
  theme(legend.title = element_blank())
```

# Notes
* Five subjects were excluded, but this was reported in the paper.
* Results from ANCOVA were not fully reported but result from ANOVA were reported and reproduced. The authors performed an ANCOVA to find no influence of chlorpromazine and then conducted and reported ANOVAs instead due to missingness on the covariate.
* Multiple follow-up ANOVAs were reported which may not be necessary and required to adjust for multiplicity, but such a correction was not performed.
* Some non-significant effects and interactions were not reported (main effect repeatability and interaction group X contour).
* The result from the ANCOVA was not clearly reported, the covariate CPZ was reported as "medication did not interact with any other variable (all ps>.6). But ANCOVA tests an overall effect of CPZ on the outcome and produces a single p-value.
* Homogeneity of variances was not met for groups.
* Independence between covariate CPZ and the IV group was met, but controls had no CPZ value.
* Homogeneity of regression slopes was met at least for the patient groups as controls had missingness.

Data was analyzed according to recommendations by Field, Miles, & Field (2012).